Human Activity Recognition in Weigtht Lifting for Coursera Practical Machine Learning Course
============================================================================================

## Introduction

  The course project [3] is about the recognition of human activity. Six participants were asked perform barbell lifts in five differnt ways (A,B,C,D,E) with sensors attached on them [4]. The goal is to identify the way they performed the exercise from the data collected.
  
## Libraries used

For the project we will have to load these three libraries:
```{r,eval=FALSE}
library("caret")
library("e1071")
library("randomForest")
```

## Data manipulation

Datasets retrieval ([5],[6]):
```{r,eval=FALSE}
tr_data<-read.csv(file="pml-training.csv",header=TRUE,sep="," )
test_data<-read.csv(file="pml-testing.csv",header=TRUE,sep="," )
```

For cross validation purposes, training data will be further sliced, 70% training dataset and 30% test dataset:
```{r,eval=FALSE}
set.seed(1111)
tr_ind <- createDataPartition(tr_data$classe, list = FALSE, p = 0.7)
tr_tr_data = tr_data[tr_ind, ]
ts_tr_data = tr_data[-tr_ind, ]
```
Numeric columns seperation:
```{r,eval=FALSE}
col_num = which(lapply(tr_tr_data, class) %in% c("numeric"))
```

The next step is to do some basic preprocessing by standardizing datasets and renaming some columns:
```{r,eval=FALSE}
preModel <- preProcess(tr_tr_data[,col_num], method = c("knnImpute"))
ptr<- cbind(tr_tr_data$classe, predict(preModel, tr_tr_data[,col_num]))
ptest<- cbind(ts_tr_data$classe, predict(preModel, ts_tr_data[, col_num]))
ptest2<-predict(preModel, test_data[, col_num])

names(ptr)[1] <- 'classe'
names(ptest)[1] <- 'classe'
names(ptest2)[1] <- 'classe'
```

## Model

We will use "random forests" algorithm to builde the model:
```{r,eval=FALSE}
model  <- randomForest(classe ~ ., ptr)
#model<-train(classe~.,data=ptr,method="rf",prox=TRUE) not enough RAM
```

Confusion Matrix on training data shows accuracy 100%:
```{r,eval=FALSE}
training_pred <- predict(rf_model, ptraining) 
print(confusionMatrix(training_pred, ptraining$classe))
```

Confusion Matrix on testing data shows accuracy 98.57%, so we don't need to prune the model:
```{r,eval=FALSE}
ptest_model<-predict(model,ptest)
print(confusionMatrix(ptest_model, ptest$classe))
```

## Results

The final step is to use the 20 test cases given (pml-testing.csv) to predict the way they perform the exercise: 
```{r,eval=FALSE}
results <- predict(rf_model, prtesting) 
results

```
The results are:

|1  |2  |3  |4  |5  |6  |7  |8  |9  |10 |11 |12 |13 |14 |15 |16 |17 |18 |19 |20 | 
|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|--:|
|B  |A  |B  |A  |A  |E  |D  |B  |A  |A  |B  |C  |B  |A  |E  |E  |A  |B  |B  |B  |


## References

[1]: Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012.

[2]: Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

[3] https://class.coursera.org/predmachlearn-002

[4] http://groupware.les.inf.puc-rio.br/har

[5] https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

[6] https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv







